---
title: "Tooth Growth"
author: "Richie Liew"
date: "Nov 16, 2016"
output:
  pdf_document: default
  html_document: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=13, fig.height=9, fig.path='Figures2/', echo=TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(ggplot2)
library(lattice)
```

### Part 2: Basic Inferential Data Analysis Instructionsless 
Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

### 1. Load the ToothGrowth data and perform some basic exploratory data analyses
```{r}
data(ToothGrowth)
str(ToothGrowth)
```

```{r}
# convert variable dose from numeric to factor
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
# look at the dataset variables after conversion
str(ToothGrowth)
```

### 2. Provide a basic summary of the data.
```{r}
head(ToothGrowth)
str(ToothGrowth)
```

```{r}
summary(ToothGrowth)
nrow(ToothGrowth)
```

```{r}
# split of cases between different dose levels and delivery methods
table(ToothGrowth$dose, ToothGrowth$supp)
```

```{r}
ggplot(aes(x=dose, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=dose))
```

```{r}
# visualization of tooth growth as function of supplement type
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp))
```

### 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
Visual examination of relationships between variables.

```{r}
xyplot(len ~ supp | dose, ToothGrowth,
       layout= c(3,1), 
       main=list(label=" Relationship between Tooth Growth, Delivery Method and Dosage",cex=0.75),
       xlab=list(label="Type of Supplement and Dosage", cex=0.75),
       ylab=list(label=" Length of Tooth Growth", cex=0.75),
       par.settings = simpleTheme(col="blue"))
```

### 4. State your conclusions and the assumptions needed for your conclusions.

1. Supplement type has no effect on tooth growth.
2. Inreasing the dose level leads to increased tooth growth.
