---
title: "Tooth Growth"
author: "Richie Liew"
date: "Nov 16, 2016"
output:
  html_document: default
  pdf_document: default
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=5, fig.height=5, fig.path='Figures2/', echo=TRUE, warning=FALSE, message=FALSE)
```

```{r, echo=FALSE}
library(ggplot2)
library(lattice)
```

### Part 2: Basic Inferential Data Analysis Instructionsless 
Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

#### 1. Load the ToothGrowth data and perform some basic exploratory data analyses
```{r}
data(ToothGrowth)
str(ToothGrowth)
```

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose) # convert from numeric to factor
str(ToothGrowth) # look at the dataset variables
```

#### 2. Provide a basic summary of the data.
```{r}
head(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)
nrow(ToothGrowth)
```

```{r}
# split of cases between different dose levels and delivery methods
table(ToothGrowth$dose, ToothGrowth$supp)
```

```{r}
# visualization of tooth growth as function of supplement type
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp))
```

### 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
Visual examination of relationships between variables.

```{r}
xyplot(len ~ supp | dose, ToothGrowth,
       layout= c(3,1), 
       main=list(label=" Relationship between Tooth Growth, Delivery Method and Dosage",cex=0.75),
       xlab=list(label="Type of Supplement and Dosage", cex=0.5),
       ylab=list(label=" Length of Tooth Growth", cex=0.5),
       par.settings = simpleTheme(col="blue"))
```

### 4. State your conclusions and the assumptions needed for your conclusions.

1. Inreasing the dose level will leads to increased in tooth growth.
2. Supplement type does not have any effect to the tooth growth.